Predicting complex diseases: performance, robustness and ..utility..(impact?)

### 2017-12-15 5-results-celiac.Rmd
save

```{r}
library(ggplot2)
myggplot <- function(..., coeff = 1) {
  bigstatsr:::MY_THEME(ggplot(...), coeff = coeff)
} 
```

@@

```{r}
# GWAS
gwas.train <- big_univLogReg(
  G, y[ind.train], ind.train = ind.train, 
  covar.train = obj.svd$u[ind.train, , drop = FALSE], 
  ncores = NCORES
)
gwas.train.gc <- snp_gc(gwas.train)
```

@@

```{r}
ggsave("figures/supp-celiac-man.png", scale = 1/90, width = 1300, height = 1000)
```

@@

>>>>

```{r}
library(tidyverse)
myggplot <- function(..., coeff = 1) {
  bigstatsr:::MY_THEME(ggplot(...), coeff = coeff)
} 
```

@@

```{r}
# GWAS
gwas.train <- big_univLogReg(
  G, y[ind.train], ind.train = ind.train, 
  covar.train = obj.svd$u[ind.train, , drop = FALSE], 
  ncores = NCORES
)
snp_qq(gwas.train) +
  coord_cartesian(ylim = c(0, 25))
gwas.train.gc <- snp_gc(gwas.train)
```

@@

@@

### ROC Curves

```{r, fig.asp=1, fig.width=8}
library(plotROC)
scores_tidy <- tibble(
  d = y[ind.test], 
  "PRS-max" =  prs[, which.max(aucs)],
  "logit-simple" = preds2,
  "logit-triple" = preds4
) %>%
  gather(key = "Method", value = "Score", -d)

bigstatsr:::MY_THEME(
  ggplot(scores_tidy, aes(d = d, m = Score, color = Method, linetype = Method)) +
    style_roc(xlab = "1 - Specificity", ylab = "Sensitivity")
) +
  geom_roc(n.cuts = 0, size = 2) +
  theme(legend.position = c(0.7, 0.3), legend.key.width = unit(4, "line")) +
  coord_equal()
```

```{r}
ggsave("figures/supp-celiac-roc.pdf", scale = 1/90, width = 700, height = 700)
```

## Top scores

First, we get the SNP with the highest ratio of cases over controls for homozygous-recessive individuals.

```{r}
props <- sapply(cols_along(G), function(j) {
  ccoj <- counts.control[, j]
  mut <- `if`(ccoj[1] < ccoj[3], 1, 3)
  `if`(counts.case[mut, j] > 100, counts.case[mut, j] / ccoj[mut], 0)
})
theone <- which.max(props)
table(Geno = G[, theone], forcats::fct_recode(as.factor(y), Control = "0", Case = "1"))
```

```{r}
# one geno
ind0 <- which(G[ind.test, theone] == 0)
(M <- length(ind0))
mean(y[ind.test[ind0]]) 
# prs max
indPRS <- order(prs[, which.max(aucs)], decreasing = TRUE)[1:M]
mean(y[ind.test[indPRS]])
# logit-simple (CMSA)
indLS <- order(preds2, decreasing = TRUE)[1:M]
mean(y[ind.test[indLS]])
# logit-triple (CMSA)
indLT <- order(preds4, decreasing = TRUE)[1:M]
mean(y[ind.test[indLT]])
```

## Stability of selection of the "logit-simple"

Train on 100 bootstrap samples of the Celiac dataset.

```{r, eval=FALSE}
if (!dir.exists("results6")) dir.create("results6")

for (i in seq_len(100)) {

  res.file <- paste0("results6/simu_", i, ".rds")
  if (file.exists(res.file)) next

  ind.train <- sort(sample(nrow(G)))

  cmsa.logit <- big_CMSA(FUN = big_spLogReg, feval = AUC,
                         X = G, y.train = y[ind.train], 
                         ind.train = ind.train, 
                         covar.train = obj.svd$u[ind.train, ],
                         alpha = 0.5, dfmax = 20e3, 
                         ncores = NCORES)

  saveRDS(cmsa.logit[cols_along(G)], file = res.file)
}
```

```{r}
results6 <- list.files("results6", full.names = TRUE) %>%
  map(~readRDS(.x)) %>%
  do.call("cbind", .)
```

```{r}
n_enter <- rowSums(results6 != 0)
mean(n_enter == 0)                ## Proportion of SNPs never entering
n_enter_once <- n_enter[n_enter != 0]
mean(n_enter_once)                ## Mean number of times a SNPs enter 
                                  ## if entering at least once
bigstatsr:::MY_THEME(
  qplot(n_enter_once, geom = "bar") +
    labs(x = "Number of times")
)
```

```{r}
ggsave("figures/supp-celiac-stabilitybar.pdf", scale = 1/90, width = 1000, height = 700)
```

```{r}
length(always_enter <- which(n_enter == ncol(results6)))
cowplot::plot_grid(
  snp_manhattan(gwas.train.gc, CHR, POS, labels = labels, 
                ind.highlight = always_enter),
  snp_manhattan(gwas.train.gc, CHR, POS, labels = labels, 
                ind.highlight = always_enter) +
    coord_cartesian(ylim = c(0, 25)) + 
    geom_hline(yintercept = -log10(5e-8), color = "red", linetype = 3),
  align = "hv", ncol = 1, labels = LETTERS[1:2], label_size = 25, scale = 0.95
)
```

```{r}
ggsave("figures/supp-celiac-man-col.png", scale = 1/90, width = 1300, height = 1000)
```

```{r}
sq_effects_always <- colSums((results6[always_enter, ])^2)
sq_effects_others <- colSums((results6[-always_enter, ])^2)
summary(100 * sq_effects_always / (sq_effects_always + sq_effects_others))
```

### 2017-12-15 paper/draft.tex
save

[MONTRER UN GRAPH EN FCT D'INCLUSION]

We also the whole Celiac dataset with the original real phenotypes. For one run, splitting the dataset in 12,000 indiduals in the training set and the remaining 3155 individuals in the test set, we compare the ``regularization paths'' of all methods (T-Trees excepted), that is to say the predictive performance as a function of the number of predictors corresponding to different inclusion thresholds of the methods.

@@

@@

>>>>

We also analyze the whole Celiac dataset with the original real phenotypes. For one run, splitting the dataset in 12,000 indiduals in the training set and the remaining 3155 individuals in the test set, we compare the ``regularization paths'' of all methods (T-Trees excepted), that is to say the predictive performance as a function of the number of predictors corresponding to different inclusion thresholds of the methods. 
We show that the regularization paths of both logistic regressions are both very smooth, reaching a maximum AUC or more than 89\% at approximately 1000 predictors (Figure \ref{fig:main-celiac-regpath}). Moreover, the CMSA procedure give AUC very close to the maximum AUC of the regularization path. For the P+T method, on the contrary, the regularization path is not smooth at all and the AUC never exceed the AUC given by the most stringent threshold [ADD VALUE OF AUC] ($p < 10^{-100}$). We also show some associated ROC curves [QUOI DIRE DE CES ROC?] (Figure \ref{fig:supp-celiac-roc}).

Then, we investigate the stability of SNP selection of the ``logit-simple''. For that, we ..replicate.. 100 times the following procedure: we sample with replacement all individuals and train the ``logit-simple'' and report the effects of each SNPs (for which a lot are 0). Over these 100 replicates, [VAL]\% of the SNPs never enter the predictions and [VAL] enter all predictions (Figure \ref{fig:supp-celiac-stabilitybar}). If we compute the sum of squared scores of the SNPs that enter every model divided by the sum of all squared scores [HAS A NAME?], we end up with a ratio of more than 99\% for all predictions, showing that most of the predictive performance is in these SNPs always entering the models. Moreover, these selected SNPs correspond to peaks of the GWAS, but also to less significant SNPs (Figure \ref{fig:supp-celiac-man-col}).

@@

\newpage
\begin{figure}[h]
\centerline{\includegraphics[width=\textwidth]{main-celiac-regpath}}
\caption{``Regularization paths'' for the P+T, ``logit-simple'' and ``logit-triple'' methods for one run on the Celiac disease, training on 12,000 individuals and projecting on the remaining 3155 individuals. For the P+T method, the dotted line represents the maximum AUC while for the logistic regressions, the dotted lines represent the value using the CMSA procedure to be free of the choice of the threshold.}
\label{fig:main-celiac-regpath}
\end{figure}

@@

\newpage
\begin{figure}[h]
\centerline{\includegraphics[width=\textwidth]{supp-celiac-roc}}
\caption{ROC Curves for the P+T, ``logit-simple'' and ``logit-triple'' methods for one run on the Celiac disease, training on 12,000 individuals and projecting on the remaining 3155 individuals.}
\label{fig:supp-celiac-roc}
\end{figure}

\newpage
\begin{figure}[h]
\centerline{\includegraphics[width=\textwidth]{supp-celiac-stabilitybar}}
\caption{Barplot showing how many times SNPs enter the predictions among predictions for 100 samples with replacement of the Celiac dataset. The bar for $0$ is not displayed for sake of visibility.}
\label{fig:supp-celiac-stabilitybar}
\end{figure}

\newpage
\begin{figure}[h]
\centerline{\includegraphics[width=\textwidth]{supp-celiac-man-col}}
\caption{Manhattan plot of the Celiac dataset. Red points are presenting SNPs that are always selected in 100 replicates of [...].}
\label{fig:supp-celiac-man-col}
\end{figure}
